---
title: "Introduction charts"
author: "Jade Talbot"
date: "08/01/2021"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction charts

How to plot charts in introduction to describe the cohort. Includes:

  + Distribution Pie Charts (Condition, Collection Type, and D'amico Risk Score)
  + Box plot of Age in each Condition
  + Age vs PSA Scatter Plot at Diagnosis with line of best fit
  + Box plot of PSA at diagnosis
  + Box plot of PSA vs Gleason Score
  
#### Distribution of Conditions

Script reference: Condition_pie_chart.R

COLOURS  
BENIGN = #694BAD  
CONTROL = #3F7EAA  
PCA = #F5C12C  

  + Load in packages
  + Load in mapping files
  + Remove remove.list (Stoller repeat list), samples collected pre-diagnosis, batch and total pools
  + Create list/dataframe of selected condition
  + Print no. of each variable

```{r}
library(tidyverse)
library(dplyr)
library(ggplot2)
library(scales)

#load in mapping file

sample_mapping_PCD01_ALL <- read.csv('../PCD01_SWATH/PCD01_general/sample_mapping_PCD01_ALL.csv')

MS_Sample_processing_LOG_Repeats <- read.csv('../PCD01_SWATH/PCD01_general/MS Sample processing LOG - Repeats.csv')

samples <- sample_mapping_PCD01_ALL %>% select('Stoller_sample_ID', Condition, "D.amico.Risk.Score", Collection_Type, 'Age_.at_collection')

remove.list <- paste(paste(MS_Sample_processing_LOG_Repeats$Stoller_Sample_ID,"$", sep = ""),sep = "|", collapse = "|")

samples <- samples[grep(remove.list, samples$'Stoller_sample_ID', invert = T), ]

samples <- samples[grep("Pre_Diagnosis", invert = T, fixed = T, samples$Collection_Type), ]

samples <- samples[grep("Control_pool|Benign_pool_|Disease_pool_|QC", samples$'Stoller_sample_ID', invert = T), ]

samples <- samples[grep("PCD01-B*", samples$'Stoller_sample_ID', invert = T), ]

samples <- samples[complete.cases(samples$Condition), ]

#Create dataframe of conditions
condition <- samples$Condition

#prints how many of each 
length(grep("Benign", condition))
length(grep("Control", condition))
length(grep("PCA", condition))
```

  + Input values into new data frame in relation to each variable, and the labels of the Pie Chart
  + Visualise on Bar Plot
  + Create blank_theme 
    + plot.title=element_text(size=) may need changing depending on amount of text in title
  + Plot Pie Chart. Change
    + ggtitle 
    + scale_fill_manual(values=c( 
    
  
```{r}
condition <- data.frame(
  group = c("Benign", "Control", "Prostate Cancer"),
  value = c(51, 167, 117),
  percentage = c("15% (51)", "50% (167)", "35% (117)")
)

#visualise data
bp <- ggplot(condition, aes(x="", y=value, fill=group))+
  geom_bar(width = 1, stat = "identity")
bp

#blanktheme
blank_theme <- theme_minimal()+
  theme(
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    panel.border = element_blank(),
    panel.grid=element_blank(),
    axis.ticks = element_blank(),
    plot.title=element_text(size=20, face="bold", hjust = 0.5),
    legend.title = element_blank()
  )

#pie chart

pie <- bp + 
  coord_polar("y", start=0, direction = -1) +
  blank_theme +
  ggtitle("Distribution of Conditions") +
  theme(axis.text.x=element_blank()) +
  scale_fill_manual(values=c("#694BAD", "#3F7EAA", "#F5C12C"), labels=c("Benign", "Control", "Prostate Cancer")) +
  geom_text(aes(label = percentage),position = position_stack(vjust = 0.5), fontface = "bold", color = "white", size=4)

pie
```

#### Comparison of age in each condition

Script reference: boxplot.R
COLOURS  
BENIGN = #694BAD  
CONTROL = #3F7EAA  
PCA = #F5C12C  

  + Load in ggplot2
  + Load in mapping file
  + Make dataframe of Condition and Age_at_Collection
  
```{r}
#box plot of age at collection

library(ggplot2)
library(dplyr)

#make data frame

sample_mapping_PCD01_ALL <- read.csv('../PCD01_SWATH/PCD01_general/sample_mapping_PCD01_ALL.csv')

samples <- sample_mapping_PCD01_ALL %>% select('Stoller_sample_ID', Condition, "D.amico.Risk.Score", Collection_Type, 'Age_.at_collection')

remove.list <- paste(paste(MS_Sample_processing_LOG_Repeats$Stoller_Sample_ID,"$", sep = ""),sep = "|", collapse = "|")

samples <- samples[grep(remove.list, samples$'Stoller_sample_ID', invert = T), ]

samples <- samples[grep("Pre_Diagnosis", invert = T, fixed = T, samples$Collection_Type), ]

samples <- samples[grep("Control_pool|Benign_pool_|Disease_pool_|QC", samples$'Stoller_sample_ID', invert = T), ]

samples <- samples[grep("PCD01-B*", samples$'Stoller_sample_ID', invert = T), ]

samples <- samples[complete.cases(samples$Condition), ]

age <- samples %>% select(Condition, Age_.at_collection)
age <- age[complete.cases(age$Condition),]

#change column headings
colnames(age)[1] <- "variable"
colnames(age)[2] <- "value"
age$value <- as.numeric(age$value)
```

  + Plot Box Plot
  
```{r}
ggplot(age, aes(x=variable , y=value, fill=variable)) + 
  geom_boxplot(aes(fill=variable)) + 
  ggtitle("Comparison of age in each condition") +
  xlab("")+ylab("Age at Collection") +
  theme(axis.title.x =element_blank(),
        axis.text.x =element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        axis.line = element_line(colour = "black"),
		aspect.ratio=1,
        plot.title = element_text(hjust = 0.5, size = 14, face = "bold")) +
scale_fill_manual(values=c("#694BAD", "#3F7EAA", "#F5C12C"), labels=c("Benign", "Control", "Prostate Cancer")) +
labs(fill = "Condition")
```

#### Comparison of Age and PSA at diagnosis and PSA vs Gleason Score

Script reference: agevspsa graph.R

  + Load in packages
  + Load in mapping file
  + Plot scatter of age vs PSA at diagnosis
  + Plot scatter of Gleason Score vs PSA at diagnosis
  + Plot box plot of Gleason Score vs PSA at diagnosis
  + Plot box plot of PSA at diagnosis
  
```{r}
library(tidyverse)
library(dplyr)
library(ggplot2)

sample_mapping_PCD01_ALL <- read.csv('../PCD01_SWATH/PCD01_general/sample_mapping_PCD01_ALL.csv')

remove.list <- paste(paste(MS_Sample_processing_LOG_Repeats$Stoller_Sample_ID,"$", sep = ""),sep = "|", collapse = "|")

samples <- sample_mapping_PCD01_ALL

samples <- samples[grep(remove.list, samples$'Stoller_sample_ID', invert = T), ]

samples <- samples[grep("Pre_Diagnosis", invert = T, fixed = T, samples$Collection_Type), ]

samples <- samples[grep("Control_pool|Benign_pool_|Disease_pool_|QC", samples$'Stoller_sample_ID', invert = T), ]

samples <- samples[grep("PCD01-B*", samples$'Stoller_sample_ID', invert = T), ]
```

  + Create dataframe, manipulate and format
  
```{r}
agevspsa <- samples %>% select('Age_.at_collection', PSA_at_Diagnosis, Gleason_Score)
```


```{r}
#convert to numeric

agevspsa$PSA_at_Diagnosis <- as.numeric(agevspsa$PSA_at_Diagnosis)
agevspsa$Age_at_Diagnosis <- as.numeric(agevspsa$'Age_.at_collection')
agevspsa$Gleason_Score <- as.numeric(agevspsa$Gleason_Score)

#log2 PSA
agevspsa$Log_PSA <- log2(agevspsa$PSA_at_Diagnosis)

#remove NA's
agevspsaNoNA <- agevspsa
agevspsaNoNA <- agevspsaNoNA[complete.cases(agevspsaNoNA$Age_at_Diagnosis), ]
agevspsaNoNA <- agevspsaNoNA[complete.cases(agevspsaNoNA$Log_PSA), ]
```

  + Plot Scatter Plot of Age at Diagnois vs PSA at Diagnosis
  + Add line of best fit, print stats
  
```{r}
#plot graph
ggplot(agevspsaNoNA, aes(x=Age_at_Diagnosis, y=Log_PSA)) +
  geom_point() +
  geom_smooth(method = lm, se = FALSE, color="red") +
  ggtitle("Comparison of Age and PSA at diagnosis") +
  xlab("Age at Diagnosis")+ylab("Log2 PSA at Diagnosis") +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        axis.line = element_line(colour = "black"),
        aspect.ratio=1,)

agevslogPSA <- lm(Age_at_Diagnosis~Log_PSA, agevspsaNoNA)
```

  + Remove NA's
  + Plot Scatter Plot of PSA at Diagnosis vs Gleason Score
  + Add line of best fit, print stats
  
```{r}
#remove NA's
agevspsaNoNA <- agevspsa
agevspsaNoNA <- agevspsaNoNA[complete.cases(agevspsaNoNA$Gleason_Score), ]
agevspsaNoNA <- agevspsaNoNA[complete.cases(agevspsaNoNA$Log_PSA), ]

#plot graph
ggplot(agevspsaNoNA, aes(x=Gleason_Score, y=Log_PSA)) +
  geom_point() +
  geom_smooth(method = lm, se = FALSE, color="red") +
  ggtitle("Comparison of Gleason Score and PSA at diagnosis") +
  xlab("Gleason Score at Diagnosis")+ylab("Log2 PSA at Diagnosis") +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        axis.line = element_line(colour = "black"),
        aspect.ratio=1,)

gleasonscorevslogPSA_scatter <- lm(Gleason_Score~Log_PSA, agevspsaNoNA)
```
 
  + Remove NA's
  + Plot Boxplots of PSA at Diagnosis vs Gleason Score
  + Add line of best fit, print stats
  
```{r}
#remove NA's
agevspsaNoNA <- agevspsa
agevspsaNoNA <- agevspsaNoNA[complete.cases(agevspsaNoNA$Gleason_Score), ]
agevspsaNoNA <- agevspsaNoNA[complete.cases(agevspsaNoNA$Log_PSA), ]
agevspsaNoNA$Gleason_Score <- as.factor(agevspsaNoNA$Gleason_Score)

ggplot(agevspsaNoNA, aes(x=(Gleason_Score) , y=Log_PSA)) + 
  geom_boxplot(aes(group=(Gleason_Score), color = (Gleason_Score)), lwd=1) + 
  geom_jitter(aes(group=(Gleason_Score), color = (Gleason_Score))) +
  ggtitle("Comparison of Gleason Score and PSA at diagnosis") +
  labs(y="Log2 PSA at Diagnosis", x="Gleason Score at Diagnosis") +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        axis.line = element_line(colour = "black"),
        aspect.ratio=1,
        legend.position = "none",
        plot.title = element_text(hjust = 0.3, size = 10, face = "bold")) +
  geom_smooth(aes(x = as.numeric(Gleason_Score, y = Log_PSA)), method = "lm", se = FALSE, color="red", formula = y~x) +
  scale_color_manual(values=c("#8890D7", "#6470DE", "#3F4ED8", "#1B2CD5", "#0D1DAD", "#0A1686", "#040832")) +
  scale_x_discrete(name = "Gleason Score at Diagnosis", limits=c("4", "5", "6", "7", "8", "9", "10"))
  
agevspsaNoNA$Gleason_Score <- as.numeric(agevspsaNoNA$Gleason_Score)
gleasonscorevslogPSA_box <- lm(Gleason_Score~Log_PSA, agevspsaNoNA)
```
```{r}
ggplot(agevspsaNoNA, aes( , y=PSA_at_Diagnosis)) +
  geom_boxplot() +
  ggtitle("PSA at Prostate Cancer diagnosis") +
  annotation_logticks(sides = "l") +
  labs(y="PSA concentration (ng/mL)") +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        axis.line = element_line(colour = "black"),
        aspect.ratio=1,
        axis.text.x=element_blank(), 
        axis.ticks.x=element_blank(), 
        legend.position = "none",
        plot.title = element_text(hjust = 0.3, size = 14, face = "bold")) +
  scale_y_log10(limits = c(1,1000))
```

  
#### Distribution of Collection Type

Script reference: Condition_pie_chart.R

COLOURS  
At_Diagnosis = #F95460  
Control = #FEB24C  
NA = #989898
Post_Diagnosis = #4CCA9D  

  + Load in packages
  + Load in mapping file
  + Create dataframe of selected condition
  + Print no. of each variable
  
```{r}
#Pie chart
  
library(tidyverse)
library(dplyr)
library(ggplot2)
library(scales)

#load in mapping file

sample_mapping_PCD01_ALL <- read.csv('../PCD01_SWATH/PCD01_general/sample_mapping_PCD01_ALL.csv')

samples <- sample_mapping_PCD01_ALL %>% select('Stoller_sample_ID', Condition, "D.amico.Risk.Score", Collection_Type, 'Age_.at_collection')

remove.list <- paste(paste(MS_Sample_processing_LOG_Repeats$Stoller_Sample_ID,"$", sep = ""),sep = "|", collapse = "|")

samples <- samples[grep(remove.list, samples$'Stoller_sample_ID', invert = T), ]

samples <- samples[grep("Pre_Diagnosis", invert = T, fixed = T, samples$Collection_Type), ]

samples <- samples[grep("Control_pool|Benign_pool_|Disease_pool_|QC", samples$'Stoller_sample_ID', invert = T), ]

samples <- samples[grep("PCD01-B*", samples$'Stoller_sample_ID', invert = T), ]

samples <- samples[complete.cases(samples$Condition), ]

#Create list of conditions
condition <- samples$Collection_Type

#prints how many of each 
length(grep("Post_Diagnosis", condition))
length(grep("At_Diagnosis", condition))
length(grep("Control", condition))
sum(is.na(samples$Collection_Type))
```

  + Input values into new data frame in relation to each variable, and the labels of the Pie Chart
  + Visualise on Bar Plot
  + Create blank_theme
    + plot.title=element_text(size=) may need changing depending on amount of text in title
  + Plot Pie Chart. Change
    + ggtitle
    + scale_fill_manual(values=c(
    
```{r}
condition <- data.frame(
  group = c("Post Diagnosis", "At Diagnosis", "Control", "NA"),
  value = c(115, 18, 165, 35),
  percentage = c("35% (115)", "5% (18)", "50% (165)", "10% (35)")
)

#visualise data
bp <- ggplot(condition, aes(x="", y=value, fill=group))+
  geom_bar(width = 1, stat = "identity")
bp
```

```{r}
#blanktheme
blank_theme <- theme_minimal()+
  theme(
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    panel.border = element_blank(),
    panel.grid=element_blank(),
    axis.ticks = element_blank(),
    plot.title=element_text(size=20, face="bold", hjust = 0.5),
    legend.title = element_blank()
  )

#pie chart
pie <- bp + 
  coord_polar("y", start=0, direction = -1) +
  blank_theme +
  ggtitle("Distribution of Collection Type") +
  theme(axis.text.x=element_blank()) +
  scale_fill_manual(values=c("#F95460", "#FEB24C","#4CCA9D", "#989898"), labels=c("At Diagnosis", "Control", "Post Diagnosis", "NA")) +
  geom_text(aes(label = percentage),position = position_stack(vjust = 0.5), fontface = "bold", color = "white", size=4)

pie
```
#### Distribution of D'amico Risk Score's

Distribution of D'amico Risk Score's
Script reference: Condition_pie_chart.R

COLOURS
LOW = #DD1C77  
INTERMEDIATE = #C994C7  
HIGH = #FDE0DD  

  + Load in packages
  + Load in mapping file
  + Print no. of each variable

```{r}
library(tidyverse)
library(dplyr)
library(ggplot2)
library(scales)

sample_mapping_PCD01_ALL <- read.csv('../PCD01_SWATH/PCD01_general/sample_mapping_PCD01_ALL.csv')

samples <- sample_mapping_PCD01_ALL %>% select('Stoller_sample_ID', Condition, "D.amico.Risk.Score", Collection_Type, 'Age_.at_collection')

remove.list <- paste(paste(MS_Sample_processing_LOG_Repeats$Stoller_Sample_ID,"$", sep = ""),sep = "|", collapse = "|")

samples <- samples[grep(remove.list, samples$'Stoller_sample_ID', invert = T), ]

samples <- samples[grep("Pre_Diagnosis", invert = T, fixed = T, samples$Collection_Type), ]

#samples <- samples[grep("Control_pool|Benign_pool_|Disease_pool_|QC", samples$'Stoller_sample_ID', invert = T), ]

samples <- samples[grep("PCD01-B*", samples$'Stoller_sample_ID', invert = T), ]

samples <- samples[complete.cases(samples$Condition), ]

#Create dataframe
two_conditions <- samples %>% select("D.amico.Risk.Score", Condition)
two_conditions <- two_conditions[ grep("PCA", two_conditions$Condition, invert = FALSE) , ]

length(grep("HIGH", two_conditions$`D.amico.Risk.Score`))
length(grep("INTERMEDIATE", two_conditions$`D.amico.Risk.Score`))
length(grep("LOW", two_conditions$`D.amico.Risk.Score`))
sum(is.na(samples$Collection_Type))

```
  + Input values into new data frame in relation to each variable, and the labels of the Pie Chart
  + Visualise on Bar Plot
  + Create blank_theme
    + Plot.title=element_text(size=) may need changing depending on amount of text in title
  + Plot Pie Chart. Change
    + ggtitle
    + scale_fill_manual(values=c(

```{r}
condition <- data.frame(
  group = c("HIGH", "INTERMEDIATE", "LOW", "NA"),
  value = c(53, 40, 15, 35),
  percentage = c("38% (53)", "28% (40)", "10% (15)", "24% (35)")
)

#visualise data
bp <- ggplot(condition, aes(x="", y=value, fill=group))+
  geom_bar(width = 1, stat = "identity")
bp

```

```{r}
#blanktheme
blank_theme <- theme_minimal()+
  theme(
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    panel.border = element_blank(),
    panel.grid=element_blank(),
    axis.ticks = element_blank(),
    plot.title=element_text(size=12, face="bold", hjust = 0.5),
    legend.title = element_blank()
  )

#pie chart
pie <- bp + 
  coord_polar("y", start=0, direction = -1) +
  blank_theme +
  ggtitle("Distribution of D'amico Risk Score's in Prostate cancer samples") +
  theme(axis.text.x=element_blank()) +
  scale_fill_manual(values=c("#8F003B", "#F2ACCA", "#DD1C77", "#989898"), labels=c("High", "Intermediate", "Low", "NA")) +
  geom_text(aes(label = percentage),position = position_stack(vjust = 0.5), fontface = "bold", color = "white", size=4)

pie
```
```